---
title: "GAE7"
author: "Jill Dixon"
format: 
  html:
    theme: custom.scss
    embed-resources: true
    code-tools: true

csl: ecology.csl

editor: visual
bibliography: references.bib
---

```{r}
#| label: SetUp
#| message: FALSE
#| echo: FALSE

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
  ) 

library(emmeans)
library(lmerTest)
library(effectsize)
library(car)
library(ggfortify)
library(patchwork)
library(broom)
library(latex2exp)       
library(kableExtra)
library(tidyverse)

options(knitr.kable.NA = '')
```

## Introduction

@Pintar.etal2018 conducted a field mesocosm experiment to explore the relative effects of predation risk and nutrient availability (a pdf of their publication is included in the GitHub repository). The predation risk manipulation was accomplished by adding either 0, 1, or 2, Golden Topminnows (*Fundulus chrysotus*) to each mesocosm. These small fish are surface-feeding predators and are known to elicit colonization and oviposition avoidance among dytiscid beetles and *Culex* mosquitoes, but not in some hydrophilid beetles. Nutrient-level treatments were created by the addition of 0, 4, or 8 g of rabbit chow (Small World Rabbit Food, Mannapro, St. Louis, MO, 40% protein).

## Methods

Please see @Pintar.etal2018 for greater detail on experimental design and data collection. In brief, the experiment was set up as a 3 X 3 factorial treatment design (9 treatment combinations). The experiment was replicated in 3 separate locations at 3 different times for a total of 9 replicates. This should have resulted in 81 observations total, however some fish died during the experiment rendering those observations not being included in the data resulting in only 69 observations.

```{r }
#| label: dataImport

df1 <- read_csv("Pintar_etal_2018.csv",
                show_col_types = FALSE)

# Block = 1 to 9 for the 9 spatiotemporal blocks (3 locations X 3 times)
# Pool = 1 to 9 for the plastic wading pools used for the experiment
# Location = 1 to 3 for the 3 fields where the pools were deployed
# Time = a to c for the 3 time periods used for each 2-week run of the exp.
# Nutrient = 1 to 3 for the 3 nutrient levels
# Fish = 1 to 3 for the 3 fish densities
# Culex_Egg_Rafts = number of egg rafts collected over the 2 weeks
# Hydrophilidae = number of hydrophilid beetles (Water Scavenger Beetles)
# Dytiscidae = number of dytiscid beetles (Predaceous Diving Beetles)

# Fish and Nutrient are provided as integers, but I wish to use them as ordered
# factors so that, by default, polynomial contrasts will be used to compute the
# ANOVAs.

df1$Nutrient <- factor(df1$Nutrient,
                       levels = c("0", "4", "8"),
                       ordered = TRUE)

df1$Fish <- factor(df1$Fish,
                   levels = c("0", "1", "2"),
                   ordered = TRUE)

df1$Block <- factor(df1$Block)

```

### Statistical Methods

```{r}
#| label: tbl-Culex
#| tbl-cap: "Effects of predation and nutrients on Culex"

CulexFactorial <- lmer(Culex_Egg_Rafts ~ Fish*Nutrient + (1|Block), df1)

CulexAnova <- tidy(anova(CulexFactorial))


```


```{r}

#| label: tbl-Hydro
#| tbl-cap: "Effects of predation and nutrients on Hydrophilidae beetles"

HydroFactorial <- lmer(Hydrophilidae ~ Fish*Nutrient + (1|Block), df1)

HydroAnova <- tidy(anova(HydroFactorial))

```


```{r}

#| label: tbl-Dytis
#| #| tbl-cap: "Effects of predation and nutrients on Dytiscidae"

DytisFactorial <- lmer(Dytiscidae ~ Fish*Nutrient + (1|Block), df1)

DytisAnova <- tidy(anova(DytisFactorial))

```


```{r}

#| label: fig-Culex
#| fig-cap: "Plot showing the effects of presence of fish predators and nutrients on Culex egg rafts."

ggplot(df1) +
  stat_summary(
    aes(x = Nutrient,
        y = Culex_Egg_Rafts,
        group = Fish,
        shape = Fish,
        color = Fish,
    ),
        position = position_dodge(
          width = 0.2)
      
    ) +
        scale_color_manual(values = c("red4", "orange", "blue3")) +
  
  theme_classic()

```


```{r}

#| label: fig-Hydro
#| fig-cap: "Plot showing the effects of presence of fish predators and nutrients on Hydrophilidae."

ggplot(df1) +
  stat_summary(
    aes(x = Nutrient,
        y = Hydrophilidae,
        group = Fish,
        shape = Fish,
        color = Fish,
    ),
        position = position_dodge(
          width = 0.2)
      
    ) +
        scale_color_manual(values = c("red4", "orange", "blue3")) +
  
  theme_classic() 

```


```{r}

#| label: fig-Dytis
#| fig-cap: "Plot showing the effects of presence of fish predators and nutrients on Dytiscidae."

ggplot(df1) +
  stat_summary(
    aes(x = Nutrient,
        y = Dytiscidae,
        group = Fish,
        shape = Fish,
        color = Fish,
    ),
        position = position_dodge(
          width = 0.2)
      
    ) +
        scale_color_manual(values = c("red4", "orange", "blue3")) +
  
  theme_classic()  

```

## Results

There was no significant relationship between predator presence and nutrient amount for Culex and Hydrophilidae (p > 0.05), but for Dytiscidae, there was a significant relationship (p < 0.01) @tbl-Culex @tbl-Hydro @tbl-Dytis.The presence of fish predators and nutrients had varying effects on the studied species. There was a significant increase in Culex egg rafts when there were no fish (p = 0.00161), as well as when nutrients increased, (p = 0.04256) @fig-Culex. Increased nutrient availability did not result in a significant increase in Hydrophilidae (p = 0.1088), and neither did the absence of predators (p = 0.0606) @fig-Hydro. Lastly, decreasing the number of fish predators and increasing nutrients both resulted in a significant increase in Dytiscidae ( p < 0.001), (p < 0.001) @fig-Dytis.

## References
